Interval Probability and Measurement for Fuzzy Quantum Theories

### 2016-08-19 prob.tex
prob.tex \subsection{Measuring Quantum Probabilities Ideally: Radon-Nikodym and Gleason's Theorem} updated...

Although specifying a probability measure for every event looks complex,
it can be simply constructed by 
\begin{equation}
\pmeas(E)=\sideset{_{\mathbb{R}}}{_{\omega\in E}}\sum\rho(\omega)\textrm{ ,}\label{eq:Radon-Nikodym}
\end{equation}
where $\rho:\Omega\rightarrow[0,1]$, $\sideset{_{\mathbb{R}}}{_{\omega\in\Omega}}\sum\rho(\omega)=1$,
and $\rho$ is called the Radon-Nikodym derivative of $\pmeas$ (with
respect to the counting measure)~\cite{Nikodym1930,Folland2013}.
The converse is also valid and called the Radon-Nikodym theorem.

\begin{thm}[Radon-Nikodym theorem for finite probability space~\cite{Kolmogorov1950,Swart2013}]
For every probability measure~$\pmeas$, there exists a unique Radon-Nikodym
derivative~$\rho$ such that equation~(\ref{eq:Radon-Nikodym})
holds.\qed\end{thm}

For example, the Radon-Nikodym derivative of $\pmeas$ in example~\ref{ex1}
is: 
\[
\begin{array}{rcl}
\rho\left(HH\right)=1/3,\quad\rho\left(HT\right)=0,\quad\rho\left(TH\right)=2/3,\quad\rho\left(TT\right)=0\textrm{ .}\end{array}
\]

@@

In contrast to real-valued probability measures, there is no easy
Radon-Nikodym theorem to simplify an interval-valued probability measure
to its Radon-Nikodym derivative. So far, the interval-valued Radon-Nikodym
Theorems are either too restricted to apply to all interval-valued
probability measures~\cite{Choquet1954,Graf1980,Goodman2013} or
corresponding an interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
to another set function~$\bar{\rho}:\events\rightarrow\bar{\mathscr{I}}$
which is not as simple as the original Radon-Nikodym derivative~\cite{GilboaSchmeidler1994,Goodman2013}.

@@

\begin{definition}[Quantum Probability Measure~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}]\label{def:QuantumProbabilitySpace}
Given a Hilbert space $\Hilb$ with its set of events~$\events$,
a \emph{quantum probability measure} is a function~$\mu:\events\rightarrow[0,1]$
such that:\footnote{It is possible to define a more general space of events consisting
of all operators~$\mathcal{A}$ on $\Hilb$ and to extend $\mu$
linearly to $\mathcal{A}$ and consider $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$
is the expectation value of $A$ which could bigger than one because
$\mu$ is linear. We does not take this approach because we want to
focus only on probability. } 
\begin{itemize}
\item $\mu(\mathbb{0})=0$, 
\item $\mu(\mathbb{1})=1$, 
\item for any projection $P$, 
\begin{equation}
\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)\textrm{ ,}\label{eq:QuantumReal-valuedProbability-Complement}
\end{equation}
and 
\item for a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\mu\left(\tensor*[_{\mathscr{O}}]{\sum}{_{i=1}^{N}}P_{i}\right)=\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\mu\left(P_{i}\right)$. 
\end{itemize}
\qed\end{definition}

@@

Moreover, the Born rule can be extended to a mixed state. When we
prepare a state, suppose the probability of preparing state~$\ket{\phi_{j}}$
is $q_{j}$, the state of the system can be expressed as a density
matrix~$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$, where $\sum_{j=1}^{N}q_{j}=1$.
Then, the quantum probability measure introduced by $\rho$ is the
combination of $\mu_{\phi_{j}}^{B}$ with respect to probability $q_{j}$~\cite{peres1995quantum,544199,RiederSvozil2007}:
\begin{eqnarray}
\mu_{\rho}^{B}\left(P\right) & = & \Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\mu_{\phi_{j}}^{B}\left(P\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}

Notice that the mixed state~$\rho$ is really a generalization of
Radon-Nikodym derivative if we pick a basis~$\left\{ \ket{\psi_{j}}\right\} _{j=1}^{d}$
such that $P=\tensor*[_{\mathscr{O}}]{\sum}{_{\ket{\psi_{j}}\in S}}\proj{\psi_{j}}$,
where $S$ is the linear subspace $P$ projects onto. By representing
the density matrix~$\rho$ in the same basis, i.e., $\rho=\sum_{i=1}^{d}\sum_{j=1}^{d}r_{ij}\ket{\psi_{i}}\bra{\psi_{j}}$,
the Born rule becomes
\begin{eqnarray}
\mu_{\rho}^{B}\left(P\right) & = & \Tr\left(\rho P\right)=\sum_{\ket{\psi_{j}}\in S}r_{jj}\textrm{ ,}\label{BornRule.mixed-1}
\end{eqnarray}
which is basically the same as equation~(\ref{eq:Radon-Nikodym}).

In this sense, the Radon-Nikodym theorem could be generalized to Gleason's
theorem~\cite{HollandJr1970,Redhead1987-REDINA,Jaeger2007}: in Hilbert
spaces of dimension $d\geq3$, given a quantum probability measure
$\mu$, there exist a mixed state~$\rho$ that induces such a measure
using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
By applying Gleason's theorem, we can extend definition~\ref{def:QuantumProbabilitySpace}
to the following corollary.

@@

%%%%%
\subsection{Measuring Quantum Probabilities}

Similar to the classical case, by applying the law of large numbers,
quantum probabilities can be estimated by relative frequencies. For
example, if we want to know the probability of the spin up in the
Stern-Gerlach experiment~\cite{Stern1988,peres1995quantum,544199,Griffiths2003},
we can put a beam of silver atoms in a highly inhomogeneous magnetic
field, and counting the number of atoms deflects up. Ideally, if the
local field of strength directs to the $z$-axis, and all particles
have the same velocity, the Stern-Gerlach experiment only produces
two spots corresponding to $\ket{0}$ and $\ket{1}$. In reality,
we need a lot of resource to keep the local field of strength directing
to the $z$-axis precisely, to keep the atoms having almost the same
velocity, and to point out the exact position each particle landed
on.

If the field of strength does not perfectly direct to the $z$-axis,
we do not really test the quantum event we want to test. This problem
will be handled later when we introduce the discrete quantum theory.
Even if the field of strength perfectly directs to the $z$-axis,
the variant velocity makes spots broader and more washed out. Together
with the precision limit of the detector, it may sometimes be hard
to decide a particle corresponding to which state\footnote{\yutsung{Add citations to support the idea... Haven't found suitable
ones...}}. Similar to Buffon's needles, this kind of fuzziness can be taken
into account by associating each quantum event with an interval-valued
probability~$\left[l,r\right]$.

\amr{preparation fuzzy, device fuzzy, Meyer~\cite{PhysRevLett.83.3751}}

@@

\subsection{Quantum Interval-valued Probability Measures}

As mentioned above, each quantum event~$P$ will be associated with
an interval-valued probability~$\left[l,r\right]$. Notice that whenever
the magnetic field of the Stern-Gerlach is fixed, i.e., an ideal measurement
is picked, measuring the spin is exactly the same as tossing a coin.
In another word, if somebody claims she is tossing a coin behind a
veil, and only shows the resulting heads or tails, we cannot distinguish
whether she has really tossed a coin, or she has run an Stern-Gerlach
experiment, and shows us the head, the tail, or the side of coin if
the silver atoms is spin up, spin down, or just hit the middle, respectively.
Therefore, we should plug in the definition of interval-valued probability
measure into the definition of quantum probability space for each
ideal measurement.

More specifically, for classical probability measures, definition~\ref{def:ClassicalProbabilitySpace}
and lemma~\ref{lem:classicalProbabilityMeasure} are almost the same
except the last condition using the equal sign~$=$ in real-valued
definition~\ref{def:QuantumProbabilitySpace} and the subset equal
sign~$\subseteq$ in interval-valued lemma~\ref{lem:classicalProbabilityMeasure}.
Therefore, for quantum probability measures, replacing the the equal
sign~$=$ in real-valued definition~\ref{def:QuantumProbabilitySpace}
by the subset equal sign~$\subseteq$ gives the only reasonable definition
for quantum interval-valued probability measures as follow.

\begin{definition}[Quantum Interval-valued Probability Measure]\label{def:QuantumInterval-valuedProbability}
Given a Hilbert space $\Hilb$ with the set of quantum events~$\events$,
and a collection of intervals~$\mathscr{I}$, a \emph{quantum $\mathscr{I}$-interval-valued
probability measure} is a function~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$, 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$, 
\item for any projection $P$, 
\begin{equation}
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=\left[1,1\right]\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(P\right)\textrm{ ,}\label{eq:QuantumInterval-valuedProbability-Complement}
\end{equation}
\item for a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\bar{\mu}\left(\sideset{_{\mathscr{O}}}{_{i=1}^{N}}\sum P_{i}\right)\subseteq\sideset{_{\mathscr{I}}}{_{i=1}^{N}}\sum\bar{\mu}\left(P_{i}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
\qed\end{definition}

>>>>

Although specifying a probability measure for every event looks complex,
it can be simply constructed by 
\begin{equation}
\pmeas(E)=\sideset{_{\mathbb{R}}}{_{\omega\in E}}\sum f(\omega)\textrm{ ,}\label{eq:Radon-Nikodym}
\end{equation}
where $f:\Omega\rightarrow[0,1]$, $\sideset{_{\mathbb{R}}}{_{\omega\in\Omega}}\sum f(\omega)=1$,
and $f$ is called the Radon-Nikodym derivative of $\pmeas$ (with
respect to the counting measure)~\cite{Nikodym1930,Folland2013}.
The converse is also valid and called the Radon-Nikodym theorem.

\begin{thm}[Radon-Nikodym theorem for finite probability space~\cite{Kolmogorov1950,Swart2013}]
For every probability measure~$\pmeas$, there exists a unique Radon-Nikodym
derivative~$f$ such that equation~(\ref{eq:Radon-Nikodym}) holds.\qed\end{thm}

For example, the Radon-Nikodym derivative of $\pmeas$ in example~\ref{ex1}
is: 
\[
\begin{array}{rcl}
f\left(HH\right)=1/3,\quad f\left(HT\right)=0,\quad f\left(TH\right)=2/3,\quad f\left(TT\right)=0\textrm{ .}\end{array}
\]

@@

In contrast to real-valued probability measures, there is no easy
Radon-Nikodym theorem to simplify an interval-valued probability measure
to its Radon-Nikodym derivative. So far, the interval-valued Radon-Nikodym
Theorems are either too restricted to apply to all interval-valued
probability measures~\cite{Choquet1954,Graf1980,Goodman2013} or
corresponding an interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
to another set function~$\bar{f}:\events\rightarrow\bar{\mathscr{I}}$
which is not as simple as the original Radon-Nikodym derivative~\cite{GilboaSchmeidler1994,Goodman2013}.

@@

\begin{definition}[Quantum Probability Measure~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}]\label{def:QuantumProbabilitySpace}
Given a Hilbert space $\Hilb$ with its set of events~$\events$,
a \emph{quantum probability measure} is a function~$\mu:\events\rightarrow[0,1]$
such that:\footnote{It is possible to define a more general space of events consisting
of all operators~$\mathcal{A}$ on $\Hilb$ and to extend $\mu$
linearly to $\mathcal{A}$ and consider $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$
is the expectation value of $A$ which could bigger than one because
$\mu$ is linear. We does not take this approach because we want to
focus only on probability. } 
\begin{itemize}
\item $\mu(\mathbb{0})=0$. 
\item $\mu(\mathbb{1})=1$. 
\item For any projection $P$, 
\begin{equation}
\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)\textrm{ .}\label{eq:QuantumReal-valuedProbability-Complement}
\end{equation}
\item for a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\mu\left(\tensor*[_{\mathscr{O}}]{\sum}{_{i=1}^{N}}P_{i}\right)=\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\mu\left(P_{i}\right)$. 
\end{itemize}
\qed\end{definition}

@@

Moreover, the Born rule can be extended to a mixed state. When we
prepare a state, suppose the probability of preparing state~$\ket{\phi_{j}}$
is $q_{j}$, the state of the system can be expressed as a density
matrix~$\rho=\tensor*[_{\mathscr{O}}]{\sum}{_{j=1}^{N}}q_{j}\proj{\phi_{j}}$,
where $\tensor*[_{\mathbb{R}}]{\sum}{_{j=1}^{N}}q_{j}=1$. Then, the
quantum probability measure introduced by $\rho$ is the combination
of $\mu_{\phi_{j}}^{B}$ with respect to probability $q_{j}$~\cite{peres1995quantum,544199,RiederSvozil2007}:
\begin{eqnarray}
\mu_{\rho}^{B}\left(P\right) & = & \Tr\left(\rho P\right)=\sideset{_{\mathbb{R}}}{_{j=1}^{N}}\sum q_{j}\mu_{\phi_{j}}^{B}\left(P\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}

\subsection{Measuring Quantum Probabilities Ideally: Radon-Nikodym and Gleason's
Theorem}

Similar to the classical case, by applying the law of large numbers,
quantum probabilities can be estimated by relative frequencies. For
example, if we want to know the probability of the spin up in the
Stern-Gerlach experiment~\cite{Stern1988,peres1995quantum,544199,Griffiths2003},
we can put a beam of silver atoms in a highly inhomogeneous magnetic
field, and counting the number of atoms deflects up. Ideally, if the
local field of strength directs to the $z$-axis, and all particles
have the same velocity, the Stern-Gerlach experiment only produces
two spots corresponding to $\ket{0}$ and $\ket{1}$. Notice that
whenever the magnetic field of the Stern-Gerlach is fixed, i.e., an
ideal measurement is picked, measuring the spin is exactly the same
as tossing a coin. In another word, if somebody claims she is tossing
a coin behind a veil, and only shows the resulting heads or tails,
we cannot distinguish whether she has really tossed a coin, or she
has run an Stern-Gerlach experiment, and shows us the head, the tail,
or the side of coin if the silver atoms is spin up, spin down, or
just hit the middle, respectively.

Mathematically, given a Hilbert space~$\Hilb$ of dimension~$d$,
we can pick a orthonormal basis~$\Omega=\left\{ \ket{\psi_{j}}\right\} _{j=1}^{d}$,
and fix a density matrix~$\rho$ representing in the same basis,
i.e., $\rho=\tensor*[_{\mathscr{O}}]{\sum}{_{i=1}^{d}}\tensor*[_{\mathscr{O}}]{\sum}{_{i=1}^{d}}r_{ij}\ket{\psi_{i}}\bra{\psi_{j}}$.
Applying the canonical inner product, we can define a Radon-Nikodym
derivative~$f_{\rho}:\Omega\rightarrow[0,1]$ by
\begin{equation}
f_{\rho}\left(\ket{\psi_{j}}\right)=\ip{\psi_{j}}{\rho\psi_{j}}=r_{jj}\textrm{ .}\label{eq:quantum-Radon-Nikodym}
\end{equation}
Furthermore, we only consider the quantum events~$\events'$ with
the form $P=\tensor*[_{\mathscr{O}}]{\sum}{_{\ket{\psi_{j}}\in E}}\proj{\psi_{j}}$,
where $E\in2^{\Omega}$. The classical probability measure~$\pmeas':2^{\Omega}\rightarrow[0,1]$
induced by the Radon-Nikodym derivative~$f_{\rho}$ is exactly the
Born rule of the mixed state~$\rho$~\cite{Maassen2010}, 
\begin{equation}
\pmeas'\left(E\right)\overset{\left(\ref{eq:Radon-Nikodym}\right)}{=}\sideset{_{\mathbb{R}}}{_{\ket{\psi_{j}}\in E}}\sum f_{\rho}\left(\ket{\psi_{j}}\right)\overset{\left(\ref{eq:quantum-Radon-Nikodym}\right)}{=}\sideset{_{\mathbb{R}}}{_{\ket{\psi_{j}}\in E}}\sum r_{jj}=\Tr\left(\rho P\right)\overset{\left(\ref{BornRule.mixed}\right)}{=}\mu_{\rho}^{B}\left(P\right)\textrm{ .}\label{BornRule.mixed-1}
\end{equation}

In this sense, the Radon-Nikodym derivative~$f_{\rho}$ is a special
case of the mixed state~$\rho$, and the Radon-Nikodym theorem is
a special case of Gleason's theorem~\cite{HollandJr1970,Redhead1987-REDINA,Jaeger2007}:
in Hilbert spaces of dimension $d\geq3$, given a quantum probability
measure $\mu$, there exist a mixed state~$\rho$ that induces such
a measure using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
By applying Gleason's theorem, we can extend definition~\ref{def:QuantumProbabilitySpace}
to the following corollary.

@@

@@

\subsection{Measuring Quantum Probabilities in Reality: Quantum Interval-valued
Probability Measures}

In reality, a Stern-Gerlach experiment requires a lot of resource
to keep the local field of strength directing to the $z$-axis precisely,
to keep the atoms having almost the same velocity, and to point out
the exact position each particle landed on. If the field of strength
does not perfectly direct to the $z$-axis, we do not really test
the quantum event we want to test. This problem will be handled later
when we introduce the discrete quantum theory. Even if the field of
strength perfectly directs to the $z$-axis, the variant velocity
makes spots broader and more washed out. Together with the precision
limit of the detector, it may sometimes be hard to decide a particle
corresponding to which state\footnote{\yutsung{Add citations to support the idea... Haven't found suitable
ones...}

\amr{preparation fuzzy, device fuzzy, Meyer~\cite{PhysRevLett.83.3751}} }.

Similar to Buffon's needles, this kind of fuzziness can be taken into
account by associating each quantum event with an interval-valued
probability~$\left[l,r\right]$. Therefore, we should plug in the
definition of interval-valued probability measure into the definition
of quantum probability space for each ideal measurement. More specifically,
for classical probability measures, definition~\ref{def:ClassicalProbabilitySpace}
and lemma~\ref{lem:classicalProbabilityMeasure} are almost the same
except the last condition using the equal sign~$=$ in real-valued
definition~\ref{def:QuantumProbabilitySpace} and the subset equal
sign~$\subseteq$ in interval-valued lemma~\ref{lem:classicalProbabilityMeasure}.
Therefore, for quantum probability measures, replacing the the equal
sign~$=$ in real-valued definition~\ref{def:QuantumProbabilitySpace}
by the subset equal sign~$\subseteq$ gives the only reasonable definition
for quantum interval-valued probability measures as follow.

\begin{definition}[Quantum Interval-valued Probability Measure]\label{def:QuantumInterval-valuedProbability}
Given a Hilbert space $\Hilb$ with the set of quantum events~$\events$,
and a collection of intervals~$\mathscr{I}$, a \emph{quantum $\mathscr{I}$-interval-valued
probability measure} is a function~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item For any projection $P$, 
\begin{equation}
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=\left[1,1\right]\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(P\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Complement}
\end{equation}
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\bar{\mu}\left(\sideset{_{\mathscr{O}}}{_{i=1}^{N}}\sum P_{i}\right)\subseteq\sideset{_{\mathscr{I}}}{_{i=1}^{N}}\sum\bar{\mu}\left(P_{i}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
\qed\end{definition}
